---
title: |
 | Pompeo College of Business 
 | Department of Economics \& Business Analytics

subtitle: 'BANL 6100: Business Analytics'
date: 'Fall 2023'
output:
  pdf_document:
    keep_tex: yes
    fig_caption: yes
    latex_engine: pdflatex
    template: unh_syllabus_template.tex
  html_document:
    df_print: paged
geometry: margin=1in

#Fill this in with any number of instructors. It uses a two column layout with a column break after each instructor.

instructor:
- name: Dr. Mehmet Balcilar
  email: MBalcilar@newhaven.edu
  telephone: (203) 479-4779 
  office: Orange Campus N125
  officehours: 1:00-3:00 p.m. MTuTh
  web: www.mbalcilar.net
 
# You can include multiple entries for a teaching assistant by filling in all these options below. Currently this is set to populate two columns on the syllabus but there is not an even column break inserted.

# teachingassistant:
#  - name: Filler
#    email: filler@fill.com
#    officehours: W T
#    office: OfficeNumber
#    sections: '301'
  
# department chair information  
chair_name: Dr. Gazi Duman
chair_email: GDuman@newhaven.edu
chair_telephone: (203) 479-4564
chair_office: Orange Campus N132A

# Course information
coursename: 'BANL 6100-08 -- Business Analytics'
semester: 'Fall 2023'
credits: '3 credits'
classroom: 'CRN 91887'
classhours: 11:15 am - 12:05 pm	MW

# running head for each page
runhead: 'Pompeo College of Business -- BANL 6100'

# whether a syllabus title should be included
syllabus_title: 'COURSE SYLLABUS'

# uncomment one of the font families and its options if available
#fontfamily: mathpazo
#fontfamily: pxfonts
fontfamily: fourier

#fontfamily: lucidabr
#fontfamilyoptions: expert,vargreek,altbullet

fontsize: 11pt
header-includes: 
  - \linespread{1.05}
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(timeDate)
library(tibble)
library(dplyr)
library(knitr)

#This code chunk sets up a table of meeting days for the course.
#You can set days of the week for course meetings (e.g. Monday and Wednesday).
#Semester start and end dates are selected and then university holidays like spring break are specified.

#Days of the week that the course meets. Must be full day name in quotes.
Days <- c("Monday", "Wednesday")

#Semester start date for classes as YYYY/M/D.
SemesterStart <- as.Date("2022/1/10")

#Semester end date for classes as YYYY/M/D.
SemesterEnd <- as.Date("2022/4/29")

#Fall break start and stop date.
FallBreak <-
  seq.Date(
    from = as.Date("2021/10/7"),
    to = as.Date("2021/10/10"),
    by = "day"
  )

#Spring break start and stop date.
SpringBreak <-
  seq.Date(
    from = as.Date("2022/02/28"),
    to = as.Date("2022/03/06"),
    by = "day"
  )

#Creates a list of days within the semester.
SemesterDays <-
  seq.Date(from = SemesterStart, to = SemesterEnd, by = "day")

#Add day before Thanksgiving.
DayBeforeThnks<-as.Date(holiday(
    year = getRmetricsOptions("currentYear"),
    Holiday = c("USThanksgivingDay")))-1

#Full list of holidays that there are no classes including breaks.
Holidays <-
  c(as.Date(holiday(
    year = getRmetricsOptions("currentYear"),
    Holiday = c("USLaborDay", "USThanksgivingDay", "USMLKingsBirthday")
  )), SpringBreak, FallBreak, DayBeforeThnks)

#Create a list of semester days excluding holidays.
SemesterDays <- SemesterDays[!SemesterDays %in% Holidays]

#Make dataframe with all semester days.
ClassDays <- data.frame(index = 1:length(SemesterDays))

#Insert dates.
ClassDays$Date <- SemesterDays

#Insert weekday designation.
ClassDays$Weekday <- weekdays(ClassDays$Date)

#Limit to the class meeting days.
MeetingDays <- ClassDays[ClassDays$Weekday %in% Days, ]

#Create numeric list of weeks.
Week <- cut.Date(MeetingDays$Date, breaks = "week")
levels(Week) <- 1:length(levels(Week))

MeetingDays$Week <- Week
MeetingDays$Class <- 1:nrow(MeetingDays)

#Replace this with c("Lecture topic 1",...,"Lecture topic N") to fill in table.

Lecture_Topic <- 1:nrow(MeetingDays)


MeetingDays <- bind_cols(MeetingDays, Lecture = Lecture_Topic)

#### Grade table ####
Grades <- tribble( ~ Grade,
                   ~ Range,
                   "A",
                   "90-100",
                   "B",
                   "80-89",
                   "C",
                   "70-79",
                   "D",
                   "60-69",
                   "F",
                   "< 60")

####Assignment weight table####
Weights <- tribble(
                    ~ Item,
                    ~ Weight,
                    "Writing scaffolding",
                    "20 %",
                    "Semester Project",
                    "25 %",
                    "Labratory Assignments",
                    "20 %",
                    "Midterm Exam",
                    "15 %",
                    "Final Exam",
                    "20 %"
                  )


```

# Course Description

This course reviews statistical concepts and methods with emphasis on data analytics and visualizations. Topics to be covered include descriptive statistics, plots and graphs for discrete and continuous data, statistical inference, regression, and selected non-parametrics including chi-square. In addition, power pivot and other Excel analytical tools will be covered. Students will obtain a solid introduction to R as a functional programming language and will be able to use Excel and R to effectively compute statistical and graphical procedures.


# Course Format 

The course will be delivered as a fully on-ground course, with every student meeting in person.

# Required Text
\emph{OpenIntro Statistics}, Christopher D. Barr, David M. Diez, and Mine Çetinkaya-Rundel, 4th Edition, 2019. This is a free eBook, available at [http://openintro.org/os/pdf](http://openintro.org/os/pdf).

New and used hard copies of the book can be purchased at [Amazon](https://www.amazon.com/OpenIntro-Statistics-Fourth-color-interior/dp/1943450226/ref=sr_1_1?crid=XKV3CDPZPZDG&keywords=OpenIntro+Statistics&qid=1692970332&sprefix=openintro+statistics%2Caps%2C84&sr=8-1).

# Required Software

All students are rewuired to use [**R**](https://www.r-project.org) and [**RStudio**](https://posit.co/products/open-source/rstudio/), a no-cost add-on coding environment for R provided by [**Posit**](https://posit.co). R serves as a statistical programming language, while R Studio operates as a third-party software user interface designed for R. These tools will find application in nearly all course assignments. R is available under the GNP General Public License, compatible with various operating systems, including Windows and Mac OS. Similarly, RStudio can be downloaded free of charge. The computer lab has already installed both R and R-Studio. You have the option to download them onto your personal computer as outlined below:

\begin{itemize}
\item[] \textbf{R:} \href{https://cran.r-project.org}{https://cran.r-project.org}
\item[] \textbf{R-Studio:} \href{https://posit.co/download/rstudio-desktop/}{https://posit.co/download/rstudio-desktop}
\end{itemize}

Posit web page for RStudio will suggest the appropriate version for your operating system.
Furthermore, the College of Business has acquired a license to locally operate the RStudio server at the University of New Haven. This server is accessible at

\begin{center}
{\bfseries \href{rstudio.newhaven.edu:8787}{rstudio.newhaven.edu:8787} }
\end{center}

Additionally, it serves as a safeguard, ensuring that all R scripts are preserved in the event that any files are inadvertently lost within your desktop directories. You will be provide with your ID and password during the first week of the class. Using the online server presents a viable alternative in case you encounter technical difficulties with R on your personal computer. The RStdio Server offers enhanced convenience, allowing access to R from home, the classroom, or the library using any browser, computer, or tablet. Moreover, it facilitates the centralization of all your coursework in one location.

Posit also offers a free cloud bases RStudio server. For more information, please refer to: [https://posit.cloud/plans/free](https://posit.cloud/plans/free).

# Additional Resources

- You can follow the steps at [Posit](https://posit.co/download/rstudio-desktop/) to install R and RStuio on your computer.
- [LinkedIn Learning](https://technology.newhaven.edu/linkedin-learning/), provided by the University of New Haven provides University faculty, staff, and students with free access to thousands of top-quality video tutorials. You can access [New Haven
LinkedIn](https://technology.newhaven.edu/linkedin-learning/) [here](https://technology.newhaven.edu/linkedin-learning/). One of the courses you may helpful is [Getting Started with R for Data Science](https://www.linkedin.com/learning/paths/getting-started-with-r-for-data-science?u=2359714), which is available [here](https://www.linkedin.com/learning/paths/getting-started-with-r-for-data-science?u=2359714)
- Youtube videos, such as MarinStatsLectures channel: [https://www.youtube.com/channel/UCaNIxVagLhqupvUiDK01Mgg](https://www.youtube.com/channel/UCaNIxVagLhqupvUiDK01Mgg)
- [R-Bloggers](https://www.r-bloggers.com) is web site to keep up-to-date with R and analytics: [https://www.r-bloggers.com](https://www.r-bloggers.com).
- [Stack Overflow](https://stackoverflow.com/questions) is public platform serving 100 million people every month about technical questions and solutions suggested by users. An answer to a question you have is. very likely to exists here or you can ask your question id not already found.

# Course Goals

The goals of this course are:

1. to help students understand basic statistics and analytics concepts, and
2. to help students learn fundamental tools that will allow them to tackle with business

analytics challenges.

# Course Learning Objectives

At the end of this course, students should be able to:

1. demonstrate basic statistical skills and data exploration,
2. conduct data analysis using appropriate statistical software, and
3. present data analysis graphically.


# Course Requirements & Assessment
Describe course assignments and assessment and where/when/if a camera will be required for each assignment and assessment.

## Expectations

Edit the statement as appropriate for this course.

<This course will require significant in-class and out-of-class commitment from each student.  The University estimates that a student should expect to spend two hours outside of class for each hour they are in a class.  (For example, a three-credit course would average six [6] hours of additional work outside of class).> 

Learning Toolbox



# Writing requirements

For this class to fulfill the (W) requirements of USA, you will be doing a semester-long research project in an area related to sedimentology. To build your skills in discipline-specific writing, we will be doing smaller writing assignments throughout the semester. The final projects will be peer-reviewed, revised, and then submitted for final grading and assessment.

# Important Dates
January 17th - Last day to drop without WD\
April 1st - Last day to drop or withdraw\
See the [Academic Calendar](https://www.southalabama.edu/academiccalendar/) for more information.

\newpage

# Course goals

1.  Observe, describe, classify, and communicate characteristics of sediments and sedimentary rocks.

2.  Develop a framework for building hypotheses about the distributions of sediments, sedimentary structures, and sedimentary environments.

3.  Communicate sedimentary geology to a variety of audiences using different genres of writing.

4.  Read and summarize scientific papers focused on sedimentary geology and related fields.

# Class textbooks

There is no required textbook for this fall, however I recommend getting these books from the library or used versions to compliment the lectures. Without a required textbook I will also be assigning technical papers to read throughout the semester. We will discuss these and you will build your technical reading skills along with your writing skills.

### For general Sedimentology and Stratigraphy
Gary Nichols, __Sedimentology and Stratigraphy__, 2nd edition, Wiley-Blackwell, ISBN: 978-1-405-13592-4

### Field focused rock ID
Dorrik A.V. Stow, __Sedimentary Rocks in the Field: A Colour Guide__, Manson Publishing Ltd, ISBN 10: 1874545693

### Improve the mechanics of your writing
William Strunk JR. and E.B. White, __The Elements of Style__, ISBN: 1594200696

### Lectures and papers
Both lecture slides and papers will be available via Canvas. Slides will be posted after lectures and assigned papers will be available in Canvas.


\newpage

```{r Schedule, echo=FALSE}
kable(MeetingDays[,c("Date",
                     "Class",
                     "Week",
                     "Lecture")
                     ],
  caption = "Course Schedule.",
  align = c("l",
            "c",
            "c",
            "l")
)
```

\newpage

# Student conduct

## Attendance policy

-   Students who complete less than 50% of assignments/exams will be assigned a failing (F) grade.

-   Review the 'Attendance and Absences Policy', in the Undergraduate and Graduate Bulletin for attendance and absences policy

## Grading

### Late policy
I will be grading assignments in batches to provide fair assessment to everyone. This means late work can be disruptive. The penalty for late work will be 20% a day after the first 24 hours late. If you have a reasonable excuse (illness, etc.) the late policy may be relaxed.

### Extra credit
Extra credit may be made available throughout the semester at my discretion.

Assignments will be returned within a week of submission.

```{r Grades_Table, echo=FALSE}
kable(Grades,
      align = c("l", "r"))
```

Assignment weighting will follow:
```{r Weights_Table, echo=FALSE}
kable(Weights,
      align = c("l", "r"))
```

\newpage
# Writing tasks
## Scaffolding tasks
Scaffolding writing tasks assigned to give you some practice and general feedback to improve your writing.
### Abstract 1

First attempt at writing an abstract using the Nature paragraph template.

### Figure Caption

Write the figure captions for several example figures using some literature examples. Learn to describe figures in text.

### Abstract 2

Write the abstract for a second paper using the feedback from the first abstract to improve your process.

### Peer Review

Peer review of preprint or publication using example template. The goal of this assignment is to focus on critiquing the logic put forward in the paper with suggestions for refining it.


## Semester project
Semester project outline of writing products.

### Project abstract

Short outline of full project highlighting the hypotheses to test, data to collect, and impact of the project. This includes no figures but could include a small table for organizing hypotheses.

### Project first draft

Full proposal draft with figures. The important aspect of this submission is assessment of the research plan and justification.

### Project peer review

Review of a peer's proposal with the goal of improving the work by increasing clarity, readability and communication.

### Project final draft

Full, final project proposal with figures, time line and clear plan for work. Focus is on clearly communicating to your peers that the work is worth doing and that you can do it.

### Project presentation

Presentation of a peer's research proposal with a focus on justifying funding their project. The goal is to convince the audience that the project can be done and will have an important impact on society.


\newpage
<!-- The sections of code below that include "child=..." are drawing from seperate .md documents. These allow for updating text in a single spot for use in multiple syllabus documents. -->
```{r Academic disruption, child='general_components/academic_disruption.md'}

```

```{r Academic honesty, child='general_components/academic_honesty.md'}

```

## Course and Teacher evaluation
Student input for the purpose of course improvement is taken very seriously and will potentially be done periodically. Please take the time to evaluate this course and the instructor, especially at the end of the semester. Evaluations will in no way affect your grade.

\newpage
# Safety

## Field Safety

### Transportation
Wear your seatbelts at all times while in moving vehicles. Do not distract the driver.

### Location awareness
Be aware of potential dangers (animals, weather, etc) and avoid them. When in a group, please think about the safety of your peers.

## Building Safety

### Fire
Random fire safety drills may occur during the course of the term. You will be expected to evacuate the building and assemble at the designated location: outside by the northeast corner of the building (follow your instructor).

### Tornado
Move away from external windows to the hallway in the basement, first, or second floor. 

### Active Shooter Incident
"Get Out" of the building as long as they can do so safely. Once you have exited the building and are at a safe location, "Call Out" to law enforcement using 911, USA Police at 460-6312, or press the Emergency button on your LiveSafe Mobile Safety App. If you can't get out, "Shelter in Place" in a secure location that is away from windows and doors, preferably a location that can be locked and barricaded.

\newpage
# On writing

## Revising written work

Nothing is perfect on the first draft. We all revise and rework our written creations for a variety of reasons. We may not communicate the point we want to get across to all readers. We may have forgotten or not known important pieces of information initially. The tone of the writing may be inappropriate for the audience. Word choice might obscure meaning. The only way we can make our writing better is by revising it based on constructive feedback from others or ourselves.

## Genres of Science writing

### Paper reviews

Critical but constructive critique of unpublished research papers. Informs the journal editor of how the work contributes to existing literature and provides guide to authors and the editor for improvement to the manuscript. Can be signed but often anonymous.

### Descriptive Reports

Descriptive reports are often made for industry or geological surveys. These describe the composition and structure of rocks in an area. Sometimes they include estimations of resource potentials or environmental impacts of industry.

### Research papers

These pieces of science writing typically test one or more closely related hypotheses in the context of existing scientific knowledge. Most have structure that is pre-defined including: abstract, introduction, methods, results, discussion, and conclusion. Figures are an integral part of most papers in the Geosciences and range from images of outcrops to abstracted summaries of observations to plots of measurements.

### Conference abstracts

These are generally a step on the path to a full research publication where results, interpretations, and context are publicly presented for the first time.

### Grant Applications

Scientific work is generally funded with public or private funds that are obtained by writing grants. Unlike research papers, grants must persuade reviewers that the research planned is important, can be done by the applicant, and can be achieved in the time frame of the funding request.

### Science Communication

Scientists also write general summaries of their work for non-specialist audiences. These can be blog posts, or lesson plans that are shared for educators. Some folks also write books for general audiences describing the processes of science.

## Common processes

For writing in the discipline, we focus on several specific questions to guide our composition:
1) Who is our audience?
2) What is the one or two sentence take-home point?
3) How quickly are we communicating the information?

Science writing, like all writing, is socially situated and caries with it the beliefs, values, and ideologies of the particular community and culture.

## Tips and tricks for good writing
Use a citation management software to organize your papers and create bibiliographies. My preference is [Zotero](https://www.zotero.org/) because it is free, open source, and works across multiple operating systems.

Write and then revise. Ask for feedback from peers or wait until you have some distance from your writing and then revise.

Prompt yourself with questions that have short, concrete answers.

\newpage
# Campus resources

```{r COVID policy, child='general_components/COVID_policy.md'}

```

```{r Writing Center, child='general_components/writing_center.md'}

```

```{r Disability Services, child='general_components/disability_services.md'}

```

```{r Food Insecurity, child='general_components/food_insecurity.md'}

```

```{r Mental Health, child='general_components/mental_health.md'}

```

```{r Disaster Plan, child='general_components/disaster_plan.md'}

```

# Final thoughts

This document is a roadmap for our semester. We learn about the Earth together and our individual experiences shape how we interpret and value data. Like all your classes, you will get out what you put into this course. Asking for help from one another and your instructors is important, don't be afraid to ask a question about something you don't know or if you want to check your knowledge about something you think you know.

__If this document is updated, a copy will be supplied to you via Canvas and changes will be announced in class.__