---
title: |
 | Pompeo College of Business 
 | Department of Economics \& Business Analytics

subtitle: 'BANL 6100: Business Analytics'
date: 'Fall 2023'
output:
  pdf_document:
    keep_tex: yes
    fig_caption: yes
    latex_engine: pdflatex
    template: unh_syllabus_template.tex
  html_document:
    df_print: paged
geometry: margin=1in

#Fill this in with any number of instructors. It uses a two column layout with a column break after each instructor.

instructor:
- name: Dr. Mehmet Balcilar
  email: MBalcilar@newhaven.edu
  telephone: (203) 479-4779 
  office: Orange Campus N125
  officehours: W 09:00 am -11:00 am
  web: www.mbalcilar.net
 
# You can include multiple entries for a teaching assistant by filling in all these options below. Currently this is set to populate two columns on the syllabus but there is not an even column break inserted.

# teachingassistant:
#  - name: Filler
#    email: filler@fill.com
#    officehours: W T
#    office: OfficeNumber
#    sections: '301'
  
# department chair information  
chair_name: Dr. Gazi Duman
chair_email: GDuman@newhaven.edu
chair_telephone: (203) 479-4564
chair_office: Orange Campus N132A

# Course information
coursename: 'BANL 6100-08 -- Business Analytics'
semester: 'Fall 2023'
credits: '3 credits'
classroom: 'BCST 307'
classhours: W 12:30 pm - 3:15 pm

# running head for each page
runhead: 'Pompeo College of Business -- BANL 6100'

# whether a syllabus title should be included
syllabus_title: 'COURSE SYLLABUS'

# uncomment one of the font families and its options if available
#fontfamily: mathpazo
#fontfamily: pxfonts
fontfamily: fourier

#fontfamily: lucidabr
#fontfamilyoptions: expert,vargreek,altbullet

fontsize: 11pt
header-includes: 
  - \linespread{1.05}
---


```{r setup, include=FALSE, error=TRUE}
knitr::opts_chunk$set(echo = TRUE)
library(timeDate)
library(tibble)
library(dplyr)
library(knitr)


#This code chunk sets up a table of meeting days for the course.
#You can set days of the week for course meetings (e.g. Monday and Wednesday).
#Semester start and end dates are selected and then university holidays like spring break are specified.

#Days of the week that the course meets. Must be full day name in quotes.
Days <- c("Wednesday")

#Semester start date for classes as YYYY/M/D.
SemesterStart <- as.Date("2023/8/28")

#Semester end date for classes as YYYY/M/D.
SemesterEnd <- as.Date("2023/12/20")

#Fall break start and stop date.
FallBreak <-
  seq.Date(
    from = as.Date("2023/10/22"),
    to = as.Date("2023/10/24"),
    by = "day"
  )

#Spring break start and stop date.
SpringBreak <-
  seq.Date(
    from = as.Date("2024/03/09"),
    to = as.Date("2024/03/17"),
    by = "day"
  )

#Fall reading days start and stop date.
FallReading <-
  seq.Date(
    from = as.Date("2023/12/12"),
    to = as.Date("2023/12/13"),
    by = "day"
  )

#Spring reading days start and stop date.
SpringReading <-
  seq.Date(
    from = as.Date("2024/04/30"),
    to = as.Date("2024/04/30"),
    by = "day"
  )

#Creates a list of days within the semester.
SemesterDays <-
  seq.Date(from = SemesterStart, to = SemesterEnd, by = "day")

#Add day before Thanksgiving.
DayBeforeThnks<-as.Date(holiday(
  year = getRmetricsOptions("currentYear"),
  Holiday = c("USThanksgivingDay")))-1

#Full list of holidays that there are no classes including breaks.
Holidays <-
  c(as.Date(holiday(
    year = getRmetricsOptions("currentYear"),
    Holiday = c("USLaborDay", "USThanksgivingDay", "USMLKingsBirthday")
  )), SpringBreak, FallBreak, FallReading, SpringReading, DayBeforeThnks)

#Create a list of semester days excluding holidays.
SemesterDays <- SemesterDays[!SemesterDays %in% Holidays]

#Make dataframe with all semester days.
ClassDays <- data.frame(index = 1:length(SemesterDays))

#Insert dates.
ClassDays$Date <- SemesterDays

#Insert weekday designation.
ClassDays$Weekday <- weekdays(ClassDays$Date)

#Limit to the class meeting days.
MeetingDays <- ClassDays[ClassDays$Weekday %in% Days, ]

#Create numeric list of weeks.
Week <- cut.Date(MeetingDays$Date, breaks = "week")
levels(Week) <- 1:length(levels(Week))

MeetingDays$Week <- Week
MeetingDays$Class <- 1:nrow(MeetingDays)

#Replace this with c("Lecture topic 1",...,"Lecture topic N") to fill in table.
#Lecture_Topic <- 1:nrow(MeetingDays)
# Make sure number elments here is equal to number of MeetingDays, i.e. nrow(MeetingDays)
Lecture_Topic <- c(
  "Module 1",
  rep("Module 2",2),
  rep("Module 3",3),
  "Module 4",
  "Midterm Exam",
  rep("Module 4",2),
  rep("Module 5",3),
  rep("Module 6",1),
  "Final Exam"
)

#check for incorrect number of topics
if (length(Lecture_Topic) != nrow(MeetingDays)) 
  stop("Number of lecture topics is not equal to number of Class Meetings")

MeetingDays <- bind_cols(MeetingDays, Lecture = Lecture_Topic)

#### Grade table ####
Grades <- tribble( ~ "Letter Grade",
                   ~ "Grades Scored Between",
                   "A+",
                   "97 to 100",
                   "A",
                   "93 to Less than 97",
                   "A-",
                   "90 to Less than 93",
                   "B+",
                   "87 to Less than 90",
                   "B",
                   "83 to Less than 87",
                   "B-",
                   "80 to Less than 83",
                   "C+",
                   "77 to Less than 80",
                   "C",
                   "73 to Less than 77",
                   "C-",
                   "70 to Less than 73",
                   "F",
                   "Less than 70"
                   )

####Assignment weight table####
Weights <- tribble(
                    ~ Item,
                    ~ Weight,
                    "Homework",
                    "25 %",
                    "Midterm Exam",
                    "35 %",
                    "Final Exam",
                    "40 %"
                  )


```

# Course Description

This course reviews statistical concepts and methods with emphasis on data analytics and visualizations. Topics to be covered include descriptive statistics, plots and graphs for discrete and continuous data, statistical inference, regression, and selected non-parametrics including chi-square. In addition, power pivot and other Excel analytical tools will be covered. Students will obtain a solid introduction to R as a functional programming language and will be able to use Excel and R to effectively compute statistical and graphical procedures.


# Course Format 

The course will be delivered as a fully on-ground course, with every student meeting in person.

# Required Text
\emph{OpenIntro Statistics}, Christopher D. Barr, David M. Diez, and Mine Çetinkaya-Rundel, 4th Edition, 2019. This is a free eBook, available at [http://openintro.org/os/pdf](http://openintro.org/os/pdf).

New and used hard copies of the book can be purchased at [Amazon](https://www.amazon.com/OpenIntro-Statistics-Fourth-color-interior/dp/1943450226/ref=sr_1_1?crid=XKV3CDPZPZDG&keywords=OpenIntro+Statistics&qid=1692970332&sprefix=openintro+statistics%2Caps%2C84&sr=8-1).

# Required Software

All students are required to use [**R**](https://www.r-project.org) and [**RStudio**](https://posit.co/products/open-source/rstudio/), a no-cost add-on coding environment for `R` provided by [**Posit**](https://posit.co). `R` serves as a statistical programming language, while `RStudio` operates as a third-party software user interface designed for `R`. These tools will find application in nearly all course assignments. `R` is available under the GNP General Public License, compatible with various operating systems, including `Windows` and `Mac OS`. Similarly, `RStudio` can be downloaded free of charge. The computer lab has already installed both `R` and `RStudio`. You have the option to download them onto your personal computer as outlined below:

\begin{itemize}
\item[] \textbf{R:} \href{https://cran.r-project.org}{https://cran.r-project.org}
\item[] \textbf{R-Studio:} \href{https://posit.co/download/rstudio-desktop/}{https://posit.co/download/rstudio-desktop}
\end{itemize}

Posit web page for `RStudio` will suggest the appropriate version for your operating system.
Furthermore, the College of Business has acquired a license to locally operate the `RStudio` server at the University of New Haven. This server is accessible at

\begin{center}
{\bfseries \href{rstudio.newhaven.edu:8787}{rstudio.newhaven.edu:8787} }
\end{center}

Additionally, it serves as a safeguard, ensuring that all `R` scripts are preserved in the event that any files are inadvertently lost within your desktop directories. You will be provide with your ID and password during the first week of the class. Using the online server presents a viable alternative in case you encounter technical difficulties with R on your personal computer. The `RStudio Server` offers enhanced convenience, allowing access to `R` from home, the classroom, or the library using any browser, computer, or tablet. Moreover, it facilitates the centralization of all your coursework in one location.

Posit also offers a free cloud bases `RStudio Server`. For more information, please refer to: [https://posit.cloud/plans/free](https://posit.cloud/plans/free).

# Additional Resources

- You can follow the steps at [Posit](https://posit.co/download/rstudio-desktop/) to install `R` and `RStudio` on your computer.
- [LinkedIn Learning](https://technology.newhaven.edu/linkedin-learning/), provided by the University of New Haven offer University faculty, staff, and students with free access to thousands of top-quality video tutorials. You can access [New Haven
LinkedIn](https://technology.newhaven.edu/linkedin-learning/) [here](https://technology.newhaven.edu/linkedin-learning/). One of the courses you may helpful is [Getting Started with R for Data Science](https://www.linkedin.com/learning/paths/getting-started-with-r-for-data-science?u=2359714), which is available [here](https://www.linkedin.com/learning/paths/getting-started-with-r-for-data-science?u=2359714)
- Youtube videos, such as MarinStatsLectures channel: [https://www.youtube.com/channel/UCaNIxVagLhqupvUiDK01Mgg](https://www.youtube.com/channel/UCaNIxVagLhqupvUiDK01Mgg)
- [R-Bloggers](https://www.r-bloggers.com) is web site to keep up-to-date with `R` and analytics: [https://www.r-bloggers.com](https://www.r-bloggers.com).
- [Stack Overflow](https://stackoverflow.com/questions) is a public platform serving 100 million people every month about technical questions and solutions suggested by users. An answer to any questions you may have is highly likely to already exist here. However, if you do not find your question already addressed, you are welcome to pose your inquiry for further assistance.

# Course Goals

The goals of this course are:

1. to help students understand basic statistics and analytics concepts, and
2. to help students learn fundamental tools that will allow them to tackle with business

analytics challenges.

# Course Learning Objectives

At the end of this course, students should be able to:

1. demonstrate basic statistical skills and data exploration,
2. conduct data analysis using appropriate statistical software, and
3. present data analysis graphically.


# Course Requirements & Assessment

## Expectations

This course will require significant in-class and out-of-class commitment from each student. The University estimates that a student should expect to spend two hours outside of class for each hour they are in a class. (For example, a three-credit course would average six [6] hours of additional work \underline{outside} of class).

This course will teach you coding in `R` language at the intermediate level. Learning coding could be challenging in the beginning. You are expected to practice coding on a continuous basis and ask help from the instructor/TAs when you encounter an issue. Learning statistical concepts and analytics will occur simultaneous to students learning `R`. It is advantageous, but not required to know basis statistical concepts.

## Exams

There will be a midterm and a final exam. Check the schedule below for the exam dates. Contact the instructor if you have any schedule conflict with another course. All exams will be taken during class session.

## Homework

There will be several homework assignments during the semester. Your answer for each homework should be uploaded to Canvas by the due date. If a submission is late, it will receive no credit. Contact the instructor if you believe that extenuating circumstances prevail but be aware that exceptions will be made only for truly exceptional situations. Extra credit homework could be assigned at the discretion of the instructor.

## Datacamp.com

It is strongly recommended that you finish a free course on statistical programming language `R` to speed up your understanding of coding in `R`. The name of this online course is `Introduction to R` and provided by [datacamp.com](datacamp.com). You receive a Certificate of Completion at the end. Also, you will get one full homework credits for completing this course. Register for [datacamp.com](datacamp.com) using [this link](https://www.datacamp.com/groups/shared_links/47a7ed52e9916a007f7d4935715bcb752027ec071fca98fb44eb10f407c7324). You may also receive an invitation from the course (arranged by the instructor). One side benefit signing up for [datacamp.com](datacamp.com) is that all other courses offered at this site will be free for you for 6 months.

## Extra Credit

All students will have periodic opportunities for extra credit throughout the semester. Exams typically include extra credit inquiries, for instance. Additionally, there will be extra credit practice assignments. For reasons of fairness, no individual opportunities for extra credit will be offered.

## Attendance and Class Participation

You are expected to participate in all class activities, and you should be prepared to contribute to the discussions and class work. This class moves at a demanding pace and requires consistent attendance to be successful. Attendance will be taken but will not be graded. It is recognized that absences may be unavoidable due to illness or family emergency. In this event, students are expected to keep up with their readings and assignments. Absences for more than two (2) weeks is detrimental to a student’s ability to keep up with the course work. In this event, the instructor’s permission is required for satisfactory completion of the course. Whether excused or not, excessive absences may result in a failing grade since these signify a failure to fulfill course requirements. Attendance will be recorded via the Qwickly Attendance application that is embedded in Canvas. Students will receive a pin code to record their attendance. If a student fails to log their attendance in Qwickly, then it is their responsibility to get with the instructor within 48 hours of the class to correct the attendance reporting issue.

# Office Hours Policy

Students have the option to either attend the designated office hours in person or initiate contact via email in order to schedule a Zoom meeting appointment. In cases where circumstances prevent direct communication with the instructor during the designated office hours, students are encouraged to arrange a mutually convenient appointment time that accommodates both the instructor and the student.  

# Course Outline/Schedule

## Topics

**Module 1 – Introduction to Business Analytics** 

  - Course Syllabus, Overview, Logistics
  - Data, Information, Knowledge, and Wisdom (DIKW)
  - `R` and `Rstudio` overview, walk through the basics
  - Basic `MS Excel` functions
  - Analyzing public companies with 10-K’s and Proxies
  - Accessing 10-Ks via `SEC.gov` and `EDGA`R
  
**Module 2 – Advanced Excel Functions and Business Analysis**

  - Pivot tables and complex `Excel` functions
  - Understanding the Business via financial statements, calculating growth rates and margins
  - Qualitative analysis based on 10-K’s and Proxies
  - Overview of the role of Statistics in Business, use case of simple linear regression model
  
**Module 3 – Data Transformation and Visualization in R**

  - Data transformation and editing with `dplyr`
  - Visualization with `ggplot2`
  - Preparing datasets for the regression model

**Module 4 – Basic Statistical Theory, Concepts and Metrics/Introduction to Statistics with R**

  - Mean, Median, Mode, Variance, Covariance, Standard Deviation and Correlation; Difference between correlation and regression
 - Skewness, how to calculate and visualize in `R`
 - Confidence intervals and _t_-test in `R` with use cases

**Module 5 – Hypothesis Testing & Regression Analysis**

  - Null and alternative hypothesis, one and two tailed tests
  - Dependent and independent variables, practicing with creating `SLR`
  - Create a linear regression model in `R` with use cases
  - Analyze the results of `SLR` and `MLR` based on the output from `R`
  - Chi-square test and `ANOVA`

**Module 6 – Data Transformation in R for Data Issues**

  - Heteroscedasticity problem; understanding the concept of heteroscedasticity
  - Using `log` to transform the data
  - Assessing the impact of data transformation on the regression model in `R`


## Course Schedule

This schedule is informational in nature and subject to change due to unforeseen circumstances, as a result of any circumstance outside the University’s control, or as other needs arise.

```{r Schedule, echo=FALSE}
kable(MeetingDays[,c("Date",
                     "Class",
                     "Week",
                     "Lecture")
                     ],
  caption = "Course Schedule.",
  align = c("l",
            "c",
            "c",
            "l")
)
```




# Grading

Grades earned are based on your performance on homework, quizzes, midterm and the final exam.

```{r Grades_Table, echo=FALSE}
kable(Grades,
      align = c("l", "r"))
```

Assignment weighting will follow:

```{r Weights_Table, echo=FALSE}
kable(Weights,
      align = c("l", "r"))
```


<!-- The sections of code below that include "child=..." are drawing from seperate .md documents. These allow for updating text in a single spot for use in multiple syllabus documents. -->

```{r Diveristy Statement, child='general_components/diversity_statement.md'}

```

```{r Reporting Bias Incidents, child='general_components/reporting_bias_incidents.md'}

```

```{r University-wide Academic Policies, child='general_components/academic_policies.md'}

```

\newpage


```{r UNIVERSITY STUDENT SUPPORT SERVICES, child='general_components/support_services.md'}

```

# Final thoughts

This document is a roadmap for our semester. We learn about the Earth together and our individual experiences shape how we interpret and value data. Like all your classes, you will get out what you put into this course. Asking for help from one another and your instructors is important, don't be afraid to ask a question about something you don't know or if you want to check your knowledge about something you think you know.

__If this document is updated, a copy will be supplied to you via Canvas and changes will be announced in class.__